\section{Text manipulation}

\subsection{Searching}
\begin{center}
\begin{longtable}{l|l}
\verb!/joe/e! & cursor set to end of match\\
\verb!3/joe/e+1! & find 3rd joe cursor set to End of match plus 1\\
\verb!/joe/s-2! & cursor set to start of match minus 2\\
\verb!/^joe.*fred.*bill/! & normal\\
\verb!/^[A-J]\+/! & search for lines beginning with one or more A-J\\
\verb!/begin\_.*end! & search over possible multiple lines\\
\verb!/fred\_s*joe/! & any whitespace including newline\\
\verb!/fred\|joe! & search for fred or joe\\
\verb!/.*fred\&.*joe! & search for fred and joe in any order\\
\verb!/\<fred\>/i! & search for fred but not alfred or frederick\\
\verb!/\<\d\d\d\d\>! & search for exactly 4 digit numbers\\
\verb!/\D\d\d\d\d\D! & search for exactly 4 digit numbers\\
\verb!/\<\d\{4}\>! & same thing\\
\verb!/\([^0-9]\|^\)%.*%! & search for absence of a digit or beginning of line\\
\verb!/^\n\{3}! & find 3 empty lines\\
\verb!/\(fred\).*\(joe\).*\2.*\1! &  using rexexp memory in a search \\
\verb!/^\([^,]*,\)\{8}! & repeating the regexp (rather than what the regexp finds) \\
\verb!/<\zs[^>]*\ze>! & search for tag contents, ignoring chevrons (\verb!:h /\zs!) \\
\verb!/<\@<=[^>]*>\@=! & search for tag contents, ignoring chevrons\\
\verb!/<\@<=\_[^>]*>\@=! & search for tags across possible multiple lines\\
\verb?/<!--\_p{-}-->? & search for multiple line comments\\
\verb!/fred\_s*joe/! & any whitespace including newline (\verb!\_!) \\
\verb!/bugs\(\_.\)*bunny! & bugs followed by bunny anywhere in file\\
\verb!:h \_! & help\\
\verb!:bufdo /searchstr/! & multiple file search (use :rewind to recommence search) \\
\verb!:bufdo %s/searchstr/&/gic! & multiple file search better but cheating (n, then a to stop) \\
\verb!?http://www.vim.org/! & search backwards for a URL without backslashing\\
\verb!/\c\v([^aeiou]&\a){4}! & search for 4 consecutive consonants
\end{longtable}
\end{center}

\subsubsection{Search for declaration of subroutine/function under cursor}
\begin{verbatim}
:nmap gx yiw/^\(sub\<bar>function\)\s\+<C-R>"<CR>
\end{verbatim}

\subsubsection{Search for visually highlighted text}
\begin{verbatim}
:vmap // y/<C-R>"<CR>
\end{verbatim}

with spec chars
\begin{verbatim}
:vmap <silent> // y/<C-R>=escape(@", '\\/.*$^~[]')<CR><CR>
\end{verbatim}

\subsection{Substitution}

\begin{center}
\begin{longtable}{l|l}
 \verb!:%s/fred/joe/igc! & general substitute command\\
 \verb!:%s/\r//g! & delete dos returns \verb!^M!\\
 \verb!:%s/\r/\r/g! & turn dos returns \verb!^M! into real returns (fixes joined lines) \\
 \verb!:%s= *\$==! & delete end of line blanks\\
 \verb!:%s= \+\$==! & same as above\\
 \verb!:%s#\s*\r\?\$##! & clean both trailing spaces AND dos returns\\
 \verb!:%s#\s*\r*\$##! & same thing deleting empty lines\\
 \verb!:%s/^\n\{3}//! & delete blocks of 3 empty lines\\
 \verb!:%s/^\n\+/\r/! & compressing empty lines\\
 \verb!:%s#<[^>]\+>##g! & delete html tags, leave text\\
 \verb!:'a,'bg/fred/s/dick/joe/igc! & VERY USEFUL\\
 \verb!:%s= [^ ]\+\$=&&=! & duplicate end column\\
 \verb!:%s= \f\+\$=&&=! & same as above\\
 \verb!:%s= \S\+\$=&&! & usually the same\\
 \verb!:s/\(.*\):\(.*\)/\2 : \1/! & reverse fields separated by :\\
 \verb!:%s/^\(.*\)\n\1\$/\1/! & delete duplicate lines\\
 \verb!:%s/^.\{-}pdf/new.pdf/! & delete to 1st pdf only\\
 \verb!:%s#\<[zy]\?tbl\_[a-z\_]\+\>#\L&#gc! & lowercase with optional leading characters\\
 \verb!:%s///! & delete possibly multi-line comments\\
 \verb!:help /\{-}! & help non-greedy\\
 \verb!:s/fred/a/g! & sub "fred" with contents of register "a"\\
 \verb!:s/fred/\=@a/g! & better alternative as register not displayed\\
 \verb!:%s/\f\+\.gif\>/\r&\r/g | v/\.gif\$/d | %s/gif/jpg/! & multiple commands on one line\\
 \verb!:%s/a/but/gie|:update|:next! & then use @: to repeat\\
 \verb!:%s/suck\|buck/loopy/gc! & ORing (must break pipe)\\
 \verb!:s/\_\_date\_\_/\=strftime("%c")/! & insert datestring\\
 \verb!:%s:\(\(\w\+\s\+\)\{2}\)str1:\1str2:! & working with columns sub any str1 in col3\\
 \verb!:%s:\(\w\+\)\(.*\s\+\)\(\w\+\)\$:\3\2\1:! & swapping first and last column (4 columns)\\
 \verb!:%s/\d\+/\=(submatch(0)-3)/! & decrement numbers by 3\\
 \verb!:g/loc\|function/s/\d/\=submatch(0)+6/! & increment numbers by 6 on certain lines only\\
 \verb!:%s#txtdev\zs\d#\=submatch(0)+1#g! & better version of above\\
 \verb!:%s/\(gg\)\@<=\d\+/\=submatch(0)+6/! & increment only numbers \verb!gg\d\d! by 6 (another way)\\
 \verb!:let i=10 | 'a,'bg/Abc/s/yy/\=i/ |let i=i+1! & convert yy to 10,11,12 etc\\
 \verb!:let i=10 | 'a,'bg/Abc/s/xx\zsyy\ze/\=i/ |let i=i+1! & convert xxyy to xx11,xx12,xx13 (more presise) \\
 \verb!:%s/"\([^.]\+\).*\zsxx/\1/! & find replacement text, use \verb!\zs! to simplify substitute\\
 \verb!:nmap <leader>z :%s#\<<c-r>=expand("<cword>")<cr>\>#! & pull word under cursor into LHS of a substitute\\
 \verb!:vmap <leader>z :<C-U>%s/\<<c-r>*\>/! & pull visually highlighted text into LHS of a substitute
\end{longtable}
\end{center}

\subsubsection{Filter all form elements into paste register}

\begin{verbatim}
:redir @*|sil exec 'g#<\(input\|select\|textarea\|/\=form\)\>#p'|redir END
:nmap ,z :redir @*<Bar>sil exec
   \ 'g@<\(input\<Bar>select\<Bar>textarea\<Bar>/\=form\)\>@p'<Bar>redir END<CR>
\end{verbatim}

\subsubsection{Substitue within substituion}
\begin{center}
\begin{longtable}{l|l}
\verb!:%s,\(all/.*\)\@<=/,_,g! & replace all / with \_ AFTER "all/"\\
\verb!:s#all/\zs.*#\=substitute(submatch(0), '/', '_', 'g')#! & same thing\\
\verb!:s#all/#&^M#|s#/#_#g|-j!! & sub by splitting line, re-joining\\
\verb!:%s/.*/\='cp '.submatch(0).' all/'.substitute(submatch(0),'/','_','g')/! & sub inside sub
\end{longtable}
\end{center}

\subsubsection{Substituting a visual area}
\begin{center}
\begin{longtable}{l|l}
 \verb!:'<,'>s/Emacs/Vim/g! & remember you DONT type the \verb!'<.'>!\\
 \verb!gv! & re-select the previous visual area (ULTRA)
\end{longtable}
\end{center}

\subsection{Global command display}
\begin{center}
\begin{longtable}{l|l}
 \verb!:g/gladiolli/#! & display with line numbers (YOU WANT THIS!)\\
 \verb!:g/fred.*joe.*dick/ & display all lines fred,joe !& dick\\
 \verb!:g/\<fred\>/! & display all lines fred but not freddy\\
 \verb!:g/^\s*\$/d! & delete all blank lines\\
 \verb?:g!/^dd/d? & delete lines not containing string\\
 \verb!:v/^dd/d! & delete lines not containing string\\
 \verb!:g/fred/,/joe/d! & not line based (very powerfull)\\
 \verb!:g/-------/.-10,.d & delete string !& 10 previous lines\\
 \verb!:g/{/ ,/}/- s/\n\+/\r/g! & delete empty lines but only between {...}\\
 \verb!:v/\S/d! & delete empty lines (both types)\\
 \verb!:v/./,/./-j! & compress empty lines\\
 \verb!:g/^\$/,/./-j! & compress empty lines\\
 \verb!:g/<input\|<form/p! & ORing\\
 \verb!:g/^/put_! & double space file (pu = put)\\
 \verb!:g/^/m0! & reverse file (m = move)\\
 \verb!:'a,'b/^/m'b! & reverse a section a to b\\
 \verb!:g/^/t.! & duplicate every line\\
 \verb!:g/fred/t\$! & copy lines matching fred to EOF\\
 \verb!:g/stage/t'a! & copy lines matching stage to marker a\\
 \verb!:g/\(^I[^^I]*\)\{80}/d! & delete all lines containing at least 80 tabs\\
 \verb!:'a,'bg/somestr/co/otherstr/! & match all lines containing "somestr" between markers a \& b\\
 \verb!:'a,'bg/str1/s/str1/&&&/|mo/str2/! & as above but also do a substitution\\
 \verb!:%norm jdd! & delete every other line\\
 \verb?:.,\$g/^\d/exe "norm! \"? & increment numbers\\
 \verb?:'a,'bg/\d\+/norm! ^A? & increment numbers\\
 \verb!:g/fred/y A! & append all lines fred to register a (empty reg a first with qaq.) \\
 \verb!:g/fred/y A | :let @*=@a! & put into paste buffer\\
 \verb!:let @a="|g/Barratt/y A |:let @*=@a! & \\
 \verb!:'a,'b g/^Error/ . w >> errors.txt! & write out to errors.txt\\
 \verb!:g/./yank|put|-1s/'/"/g|s/.*/Print '&'/! & duplicate every line in a file wrap a print " around each duplicate\\
 \verb!:g/^MARK\$/r tmp.ex | -d! & replace string with contents of a file, -d deletes the "mark"\\
 \verb!:g/<pattern>/z#.5! & display with context\\
 \verb!:g/<pattern>/z#.5|echo "=========="! & display beautifully\\
 \verb!:g/|/norm 2f|r*! & replace 2nd | with a star\\
 \verb?:nmap <f3>  :redir @a<cr>:g//<cr>:redir END<cr>:new<cr>:put! a<cr><cr>? & send output of previous global command to a new window
\end{longtable}
\end{center}

\subsection{Global combined with substitute (power editing)}
\begin{center}
\begin{longtable}{l|l}
 \verb!:'a,'bg/fred/s/joe/susan/gic! & can use memory to extend matching\\
 \verb!:g/fred/,/joe/s/fred/joe/gic! & non-line based (ultra)\\
 \verb!:/fred/;/joe/-2,/sid/+3s/sally/alley/gIC! & find fred before beginning search for joe
\end{longtable}
\end{center}

\subsection{Changing case}
\begin{center}
\begin{longtable}{l|l}
\verb!guu! & lowercase line\\
\verb!gUU! & uppercase line\\
\verb!Vu! & lowercase line\\
\verb!VU! & uppercase line\\
\verb!g$\sim$$\sim$! & flip case line\\
\verb!vEU! & upper case word\\
\verb!vE$\sim$! & flip case word\\
\verb!ggguG! & lowercase entire file\\
\verb!vmap ,c :s/\<\(.\)\(\k*\)\>/\u\1\L\2/g! & titlise visually selected text (map for .vimrc)\\
\verb=:%s/[.!?]\_s\+\a/\U&\E/g= & uppercase first letter of sentences\\
\verb!g<C-G>! & count words in text file
\end{longtable}
\end{center}

\subsection{Reformatting text}
\begin{center}
\begin{longtable}{l|l}
 \verb!gq}! & format a paragraph\\
 \verb!gqap! & format a paragraph\\
 \verb!ggVGgq! & reformat entire file\\
 \verb!Vgq! & current line\\
 \verb!:s/.\{,69\};\s*\|.\{,69\}\s\+/&\r/g! & break lines at 70 chars, if possible after a ';'\\
\end{longtable}
\end{center}

\subsection{Deletion without destroying buffer}
\begin{center}
\begin{longtable}{l|l}
 \verb!"_d! & what you've ALWAYS wanted\\
 \verb!"_dw! & delete word (use blackhole)
\end{longtable}
\end{center}

\subsection{Essential}
\begin{center}
\begin{longtable}{l|l}
\verb!* # g* g#! & find word under cursor () (forwards/backwards)\\
\verb!%! & match brackets and tags \{\}, [], (), etc. \\
\verb!.! & repeat last modification \\
\verb!@:! & repeat last : command (then @@)\\
\verb!<C-N><C-P>! & word completion in insert mode\\
\verb!<C-X><C-L>! & line complete SUPER USEFUL\\
\verb!/<C-R><C-W>! & pull onto search/command line\\
\verb!/<C-R><C-A>! & pull onto search/command line\\
\verb!:set ignorecase! & you nearly always want this\\
\verb!:syntax on! & colour syntax in perl, HTML, PHP etc. \\
\verb!:h regexp! & list all help topics containing regexp (TAB to step through list)\\
\verb!:nmap ,s :source \$VIM/_vimrc! & read from vimrc\\
\verb!:nmap ,v :e \$VIM/_vimrc! & open and edit local vimrc \\
\verb!:vmap sb "zdi<b><C-R>z</b><ESC>! & wrap \verb!<b></b>! around VISUALLY selected text\\
\verb!:vmap st "zdi<?= <C-R>z ?><ESC>! & wrap \verb!<?= ?>! around VISUALLY selected text
\end{longtable}
\end{center}
